---
title: "Simulation Exercise and Basic Data Analysis"
author: "Jean-Paul Courneya"
date: "8/6/2017"
output:
  pdf_document: default
  html_document: default
---
#This class project consists of two parts:
#### 1. A simulation exercise.
#### 2. Basic inferential data analysis.

First you will set a working directory and load dependencies
```{r, eval=FALSE}
setwd()
library(knitr)
library(ggplot2)
```
***
#Part 1: Simulation Exercise 


##### In this part of the project we are investigating the exponential distribution in R and comparing it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. Set lambda = s for all of the simulations. 

##Investigating the distribution of averages of 40 exponentials.

###Illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponentials. 

Use set.seed so that the simulation can be repeated. Setting the seed provides the same starting value to the random number generating function.
```{r}
set.seed(420)
```

#####Start by taking a look at the distribution of a sample of 40 random exponentials
```{r}
hist(rexp(40,0.2))
```
#####Next we will evaluate the mean of the sample data set.
```{r}
mean(rexp(40,0.2))
```
