---
title: "Basic inferential data analysis - Project Stats-Inf"
author: "Jean-Paul Courneya"
date: "9/5/2017"
output:
  pdf_document: default
  html_document: default
---
#Part 2: Basic inferential data analysis.

```{r}
library(knitr)
library(ggplot2)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

####Load the ToothGrowth data and perform some basic exploratory data analyses

```{r}
data("ToothGrowth")
str(ToothGrowth)
```
The data for the dose is currently numeric. Since the doses are factors they will be changed to factors. 

```{r}
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
str(ToothGrowth)
```

\pagebreak

####What does the Tooth Growth data look like?
```{r}
qplot(dose, len, data = ToothGrowth, facets = supp ~ ., color = dose, xlab = "Dose (mg/day)", ylab = "Tooth Length (mm)", main = "Exploratory plot of dose vs tooth length\n for Guinea Pigs given OJ vs VitaminC")
```
\pagebreak
```{r}
f <- ggplot(ToothGrowth, aes(x=dose, y=len, fill=supp))
f + geom_boxplot() + ggtitle("Guinea Pig Tooth\ngrowth after consumption\nof OJ or Vitamin C ") +
  labs(x="Dose (mg/day)",y="Length (mm)") + 
  theme(plot.title = element_text(family = "serif", color="#666666", face="bold", size=16, hjust=0.5)) +
  theme(axis.title = element_text(family = "serif", color="#666666", face="bold", size=11))
```

***

###A basic summary of the data

```{r}
summary(ToothGrowth)
```

\pagebreak

#Compare tooth growth by supp and dose using confidence intervals and/or hypothesis tests

####The overall goal of understanding this data is achieved by determining if vitamin C affects tooth growth and if it matters how the vitamin C is delivered. 

####A t.test can be used to compare the tooth len growth between supplements (VitaminC or OJ) and each dose.

#####Tooth len growth at a dose of 0.5 mg/day (VitaminC or OJ)
Comparison at (p < 0.05) 
```{r}
t.test(ToothGrowth$len[ToothGrowth$supp == "OJ" & ToothGrowth$dose == 0.5], ToothGrowth$len[ToothGrowth$supp == "VC" & ToothGrowth$dose == 0.5])
````

####OJ has a higher effect on tooth growth at 0.5 mg/day

#####Tooth len growth at a dose of 1.0 mg/day (VitaminC or OJ)
Comparison at (p < 0.05) 
```{r}
t.test(ToothGrowth$len[ToothGrowth$supp == "OJ" & ToothGrowth$dose == 1.0], ToothGrowth$len[ToothGrowth$supp == "VC" & ToothGrowth$dose == 1.0])
````

####OJ has a higher effect on tooth growth at 1.0 mg/day

\pagebreak

#####Tooth len growth at a dose of 2.0 mg/day (VitaminC or OJ)
Comparison at (p < 0.05)
```{r}
t.test(ToothGrowth$len[ToothGrowth$supp == "OJ" & ToothGrowth$dose == 2.0], ToothGrowth$len[ToothGrowth$supp == "VC" & ToothGrowth$dose == 2.0])
````

####There is no difference between OJ and VitaminC on tooth growth at 2.0 mg/day

####Ive just demonstrated that there is a difference in tooth growth between supplements. at the low doses (0.5 mg/day, 1.0 mg/day) OJ has more of an effect on Tooth growth  but not at 2.0 mg/day

***

####Next I will compare tooth growth based on dose of supplement (OJ or VC). 

#####First I will compare 0.5 mg/day vs 1 of OJ
Comparison at (p < 0.05).
```{r}
t.test(ToothGrowth$len[ToothGrowth$supp == "OJ" & ToothGrowth$dose == 0.5], ToothGrowth$len[ToothGrowth$supp == "OJ" & ToothGrowth$dose == 1.0])
````

There is a significant difference. 

#####Here is the comparison 0.5 mg/day vs 2 of OJ 
Comparison at (p < 0.05).
```{r}
t.test(ToothGrowth$len[ToothGrowth$supp == "OJ" & ToothGrowth$dose == 0.5], ToothGrowth$len[ToothGrowth$supp == "OJ" & ToothGrowth$dose == 2.0])
````

There is a significant difference

#####Here is the comparison 1.0 mg/day vs 2 of OJ 
Comparison at (p < 0.05).
```{r}
t.test(ToothGrowth$len[ToothGrowth$supp == "OJ" & ToothGrowth$dose == 1.0], ToothGrowth$len[ToothGrowth$supp == "OJ" & ToothGrowth$dose == 2.0])
````
The difference is significant but not the same maginitude of effect as between 0.5 mg/day and (1 or 2)

####Next I will compare the effect of Vitamin C

#####First I will compare 0.5 mg/day vs 1 of VC
Comparison at (p < 0.05).
```{r}
t.test(ToothGrowth$len[ToothGrowth$supp == "VC" & ToothGrowth$dose == 0.5], ToothGrowth$len[ToothGrowth$supp == "VC" & ToothGrowth$dose == 1.0])
````

There is a significant difference. 

#####Here is the comparison 0.5 mg/day vs 2 of VC 
Comparison at (p < 0.05).
```{r}
t.test(ToothGrowth$len[ToothGrowth$supp == "VC" & ToothGrowth$dose == 0.5], ToothGrowth$len[ToothGrowth$supp == "VC" & ToothGrowth$dose == 2.0])
````

There is a significant difference

#####Here is the comparison 1.0 mg/day vs 2 of VC 
Comparison at (p < 0.05).
```{r}
t.test(ToothGrowth$len[ToothGrowth$supp == "VC" & ToothGrowth$dose == 1.0], ToothGrowth$len[ToothGrowth$supp == "VC" & ToothGrowth$dose == 2.0])
````

I have just demonstrated that not only is there a difference between the supplements effect on tooth growth by dose but that also tooth growth is significantly affected by OJ and Vitamin C. There is a maximum benefit which starts to level off at a dose of 2.0 mg/day since we see the percent difference between a dose of 1 and 2 and effect of teeth growth is decreasing regardless of supplement given.

Overall Guinea Pig tooth growth is most affected by OJ supplement.There is a noticible effect between doses of Vitamin C however within doses. 